Tweet Classification: Filtering \\ of Twitter for Company-relevant Tweets

### 2012-12-07 paper/paper.tex
paper updates

[INSERT LIST OF KEYWORDS]...\\
"apple", "itunes", "ipad", "ipod", "mac", "ios", "iphone", "AAPL", "cupertino", "safari", "ilife", "iwork", "garageband", "ibook", "powerbook", "itouch", "app store", "macworld", "facetime", "icloud", "mobileme", "siri", "imovie", "iphoto", "quicktime", "logic pro", "think different" \\

@@

To evaluate this large set of results, we have chosen three relevant evaluation metrics. The first is precision, or the ratio of correctly labeled tweet instances to the number of instances assigned a particular label (`apple' or `none' in our case). Higher values for this metric mean fewer false positives for the desired company, which is the overall goal in this classification task. The second metric was recall, or the ratio of correctly classified instances out of the full set of instances which should have the label in question. This metric is less relevant to our purpose, but still interesting. The sheer scale of Twitter means missing even half of the tweets mentioning a company and mislabeling them as about `none' is of little consequence. Finally we use the F1 metric, a standard combination of precision and recall that is useful in determining overall if some pipeline combination is better than another in both precision and recall.

@@

Our desired result was a high precision classifier that could be useful gathering a large number of tweets mentioning a company or its products. Looking at an overview of the results in Figure~\ref{fig:chart_all}, we can see all Pipeline combinations had over 70\% of ``apple'' labelings correct with only 50 training tweet instances. However, given that a baseline strategy of labeling every tweet as ``apple'' would be correct 64\% of the time, this is only an improvement of approximately 9\% in the worst case. The more relevant and interesting result is the 43\% improvement from the baseline seen with our custom Twitter Tokenizer and all other Pipes like Lowsercasing, Stemming and Bigrams trained on the maximum of 1500 instances. This relatively large improvement indicates not only is the tokenization strategy used important, but also that we were relatively successful in creating an effective tokenizer that can be further augmented by standard natural language processing techniques like stemming, removing stop words, applying n-grams, or simply lowercasing the text. See in Table \ref{tab:top_pipelines} below that all of top five Pipeline combinations include our twitter tokenizer. In fact, one must go down to the 24th and 33rd ranked combinations to find ones that include the Number-Digit and Whitespace based tokenizers respectively.

\begin{table}[h]
\centering
\small
\begin{tabular}{|c|l|r|}
	\hline
	\textit{Rank} & \textit{Pipeline Description} & \textit{Precision} \\
	& & \textit{for Apple} \\ \hline
	1 & Twitter Tokenizer, Lowercased, & \\ & Stemmed, Stopwords Removed, Bigrams & 92.2\% \\ \hline
	% TODO FIIIIIIIX
	2 & Twitter Tokenizer, Lowercased, & \\ & Stopwords Removed, Bigrams & 91.6\% \\ \hline
	3 & Twitter Tokenizer, Stopwords Removed, & \\ & Stemmed, Bigrams & 91.6\% \\ \hline
	4 & Twitter Tokenizer, Stemmed, Bigrams & 91.4\% \\ \hline
	5 & Twitter Tokenizer, Stopwords Removed, & \\ & Bigrams & 91.3\% \\ \hline
	\hline
	24 & Number-Digit Tokenizer, & \\ & Remove Stopwords, Bigrams & 88.2\% \\ \hline
	\hline
	33 & Whitespace Tokenizer, Stemmed, & \\ & Remove Stopwords, Bigrams & 87.7\% \\ \hline

\end{tabular}
\caption{Top performing Pipeline combinations for 1500 training instances}
\label{tab:top_pipelines}
\end{table}

Intro

Using whitespace-based tokenization with a Naive Bayes classifier, we achieve a baseline accuracy of 76\% %correct this exactly!
compared to the prior probability of tweet about Apple of 64\%. For simplicity of analysis, we compared the results of every permutation of our pipeline steps, training on 100 through 1800 tweets in increments of 100, in order to find the combinations that exceed or violate our expectations. We found that our Twitter-aware tokenizer outperformed whitespace tokenization by [THIS AMOUNT]. Applying stemming and stop word removal to both pipelines increased the accuracy for almost every case, though widening the gap.

@@

\subsection{Pipe Effectiveness Breakdown}

\textit{Stopword removal}: \\
~\\
\textit{Bigrams}: \\
~\\
\textit{URL Replacement}: \\
~\\
\textit{Spell checking}: Correcting misspelled words provided no statistically significant change in our results. \\

\subsection{}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{chart_twitterVSwhitespace}
\caption{Accuracy With Stemming and Stopword Removal}
\label{fig:chart_twitterwhitespace}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{chart_twitterVSwhitespace_detailed}
\caption{Accuracy For Twitter and Whitespace Tokenizers}
\label{fig:chart_twitterwhitespace_detailed}
\end{figure}

\begin{table*}[ht]
\centering
%\small
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|c|}{\textbf{Twitter-aware tokenizer}} & \multicolumn{4}{|c|}{\textbf{alphanumeric\_ tokenizer}} & \multicolumn{4}{|c|}{\textbf{whitespace tokenizer}} \\
\hline
\multicolumn{2}{|c|}{\textbf{no removal}} & \multicolumn{2}{|c|}{\textbf{stopword removal}} &\multicolumn{2}{|c|}{\textbf{no removal}} & \multicolumn{2}{|c|}{\textbf{stopword removal}} &\multicolumn{2}{|c|}{\textbf{no removal}} & \multicolumn{2}{|c|}{\textbf{stopword removal}}\\
\hline
\textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} \\ \hline
  Apple  & i     & Apple      & apple        & t        & t     & Apple  & http   & Apple  & apple & Apple   & apple  \\
  iphone & apple & iphone     & Apple        & co       & co    & http   & apple  & to     & I     & iPhone  & Apple  \\
  i      & the   & ...        & ...          & Apple    & http  & iPhone & Apple  & I      & the   & -       & -      \\
  the    & Apple & ipad       & mac          & http     & I     & iPad   & Mac    & the    & Apple & RT      & Mac    \\
  to     & and   & ipod       & juice        & I        & apple & RT     & ED     & my     & to    & iPad    & RT     \\ 
  my     & to    & rt         & rt           & to       & Apple & apple  & juice  & a      & and   & iPod    & juice  \\
  a      & a     & \#apple    & post         & the      & the   & iPod   & RT     & iPhone & a     & 5       & \&     \\
  ...    & ...   & 5          & new\_post    & iPhone   & to    & 5      & iPhone & for    & for   & iphone  & iPhone \\
  on     & for   & iphone\_5  & iphone       & my       & and   & iphone & post   & on     & -     & apple   & post:  \\
  for    & new   & my\_iphone & apple\_juice & a        & a     & ED     & A0     & and    & my    & \#apple & I'm    \\ \hline
\end{tabular}
\caption{Ten most common features for various pipelines.}
\end{table*}

@@

>>>>

\begin{tabular}{l l l}
apple & itunes & ipad \\
ipod & mac & ios \\
iphone & AAPL & cupertino \\
safari&ilife&iwork \\
garageband&ibook&powerbook \\
itouch&app store&macworld \\
facetime&icloud&mobileme \\
siri&imovie&iphoto \\
quicktime&logic pro&think different
\end{tabular} \\
~\\

@@

To evaluate this large set of results, we have chosen three relevant evaluation metrics. The first is precision, or the ratio of correctly labeled tweet instances to the number of instances assigned a particular label (``apple'' or ``none'' in our case). Higher values for this metric mean fewer false positives for the desired company, which is the overall goal in this classification task. The second metric was recall, or the ratio of correctly classified instances out of the full set of instances which should have the label in question. This metric is less relevant to our purpose, but still interesting. The sheer scale of Twitter means missing even half of the tweets mentioning a company and mislabeling them as about `none' is of little consequence. Finally we use the F1 metric, a standard combination of precision and recall that is useful in determining overall if some pipeline combination is better than another in both precision and recall.

@@

Our desired result was a high precision classifier that could be useful gathering a large number of tweets mentioning a company or its products. Looking at an overview of the results in Figure~\ref{fig:chart_all}, we can see all Pipeline combinations had over 70\% of ``apple'' labelings correct with only 50 training tweet instances. However, given that a baseline strategy of labeling every tweet as ``apple'' would be correct 64\% of the time, this is only an improvement of approximately 9\% in the worst case. The more relevant and interesting result is the 43\% improvement from the baseline seen with our custom Twitter Tokenizer and all other Pipes like Lowsercasing, Stemming and Bigrams trained on the maximum of 1500 instances. This relatively large improvement indicates not only is the tokenization strategy used important, but also that we were relatively successful in creating an effective tokenizer that can be further augmented by standard natural language processing techniques like stemming, removing stop words, applying n-grams, or simply lowercasing the text. We can see in Table \ref{tab:top_pipelines} below that all of top five Pipeline combinations include our twitter tokenizer. In fact, one must go down to the 24th and 33rd ranked combinations to find ones that include the Number-Digit and Whitespace based tokenizers respectively.

\begin{table}[h]
\centering
\small
\begin{tabular}{|c|l|r|}
	\hline
	\textit{Rank} & \textit{Pipeline Description} & \textit{Precision} \\
	& & \textit{for Apple} \\ \hline
	1 & Twitter Tokenizer, Lowercased, & \\ & Stemmed, Stopwords Removed, Bigrams & 92.2\% \\ \hline
	2 & Twitter Tokenizer, Lowercased, & \\ & Stopwords Removed, Bigrams & 91.6\% \\ \hline
	3 & Twitter Tokenizer, Stopwords Removed, & \\ & Stemmed, Bigrams & 91.6\% \\ \hline
	4 & Twitter Tokenizer, Stemmed, Bigrams & 91.4\% \\ \hline
	5 & Twitter Tokenizer, Stopwords Removed, & \\ & Bigrams & 91.3\% \\ \hline
	\hline
	24 & Alphanumeric Tokenizer, & \\ & Remove Stopwords, Bigrams & 88.2\% \\ \hline
	\hline
	33 & Whitespace Tokenizer, Stemmed, & \\ & Remove Stopwords, Bigrams & 87.7\% \\ \hline
\end{tabular}
\caption{Top performing Pipeline combinations for 1500 training instances}
\label{tab:top_pipelines}
\end{table}

@@

\subsection{Pipe Effectiveness}

\textit{Stemming}: \\
~\\
\textit{Stopword removal}: \\
~\\
\textit{Bigrams}: \\
~\\
\textit{URL Replacement}: \\
~\\
\textit{Spell checking}: Correcting misspelled words provided no statistically significant change in our results. \\

\subsection{Tokenizers}

\begin{figure}[h]
\centering
\includegraphics[]{img/tokenizers}
\caption{Precision of main tokenizers}
\label{fig:tokenizer_precision}
\end{figure}

We found that our Twitter-aware tokenizer outperformed whitespace tokenization by [THIS AMOUNT] on average across the . Applying stemming and stop word removal to both pipelines increased the accuracy for almost every case, though widening the gap.

% \begin{figure}[ht]
% \centering
% \includegraphics[width=\linewidth]{chart_twitterVSwhitespace}
% \caption{Accuracy With Stemming and Stopword Removal}
% \label{fig:chart_twitterwhitespace}
% \end{figure}

\begin{table*}[ht]
\centering
%\small
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|c|}{\textbf{Twitter-aware tokenizer}} & \multicolumn{4}{|c|}{\textbf{alphanumeric\_ tokenizer}} & \multicolumn{4}{|c|}{\textbf{whitespace tokenizer}} \\
\hline
\multicolumn{2}{|c|}{\textbf{no removal}} & \multicolumn{2}{|c|}{\textbf{stopword removal}} &\multicolumn{2}{|c|}{\textbf{no removal}} & \multicolumn{2}{|c|}{\textbf{stopword removal}} &\multicolumn{2}{|c|}{\textbf{no removal}} & \multicolumn{2}{|c|}{\textbf{stopword removal}}\\
\hline
\textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} & \textbf{apple} & \textbf{none} \\ \hline
  Apple  & i     & Apple      & apple        & t        & t     & Apple  & http   & Apple  & apple & Apple   & apple  \\
  iphone & apple & iphone     & Apple        & co       & co    & http   & apple  & to     & I     & iPhone  & Apple  \\
  i      & the   & ...        & ...          & Apple    & http  & iPhone & Apple  & I      & the   & -       & -      \\
  the    & Apple & ipad       & mac          & http     & I     & iPad   & Mac    & the    & Apple & RT      & Mac    \\
  to     & and   & ipod       & juice        & I        & apple & RT     & ED     & my     & to    & iPad    & RT     \\ 
  my     & to    & rt         & rt           & to       & Apple & apple  & juice  & a      & and   & iPod    & juice  \\
  a      & a     & \#apple    & post         & the      & the   & iPod   & RT     & iPhone & a     & 5       & \&     \\
  ...    & ...   & 5          & new\_post    & iPhone   & to    & 5      & iPhone & for    & for   & iphone  & iPhone \\
  on     & for   & iphone\_5  & iphone       & my       & and   & iphone & post   & on     & -     & apple   & post:  \\
  for    & new   & my\_iphone & apple\_juice & a        & a     & ED     & A0     & and    & my    & \#apple & I'm    \\ \hline

\end{tabular}
\caption{Ten most common features for various pipelines.}
\end{table*}

@@

\section{Team Contributions}
Specific team member contributions followed from what we originally stated when the project began. Brandon was in charge of writing some of the Java classes that would be used by the Mallet solftware to push and pull tweets to and from the server where they were stored. He also wrote and researched about different Pipes that could be used by Mallet to transform the tweets being classified. Finally, Brandon had a hand in defining the project structure, and helped edit the project's paper and website. Curtis, being the resident expert in the Mallet software, was tasked with integrating it into the project. He also wrote the majority of the paper and did extensive work, writing Pipes to manipulate tweet data. Dan initally set up a server to pull tweets from Twitter so we would have a large data set to work with. From there he hand built a testing suite to analyze the impact each of the Pipes had on classification. Because of this, Dan also compiled the majority of the statistical results that came from running his test suite.

We, Brandon Maxwell, Curtis Ullerich, and Daniel Stiner, hereby concur with the contents of this report, in the knowledge that it is an accurate representation of the work we did and the results we found, for our term project.
